Motmot: A Distributed Chat Protocol

### 2012-05-04 paper/paper.tex
More design

\subsection{Design}

@@

This simplification to Paxos yields many benefits.  For instance, clients may
still choose to keep a record of chats in a local file, but by disallowing
protocol rejoins, this persistent storage does not play any role in the safety
or progress of \libmotmot's Paxos implementation.  We can also avoid the need
for complicated and high-bandwidth recovery mechanisms.

\libmotmot also implements a number of out-of-band protocols which support
the core Paxos algorithm.  We will now describe each of these side protocols.

@@

@@

\thebibliography{2}{
  \bibitem{paxlive}
    T. Chandra, R. Griesemer, and J. Redstone.
    Paxos made live---an engineering perspective.
    In \emph{Proc. 26th Symp. on Principles of Distributed Computing}, pp. 398-407, Aug. 2007.

@@

  \bibitem{rfc}
    P. Resnick, ed.
    Internet Message Format, RFC 2822.
    April 2001.
}

>>>>

\subsection{Simplification}

@@

This simplification to Paxos yields many benefits.  For instance, clients may
still choose to keep a record of chats in a local file, but by disallowing
protocol rejoins, this persistent storage does not play any role in the safety
or progress of \libmotmot's Paxos implementation.  We can also avoid the need
for complicated and high-bandwidth recovery mechanisms.  In \libmotmot's Paxos,
as in the final protocol presented in the original Paxos paper, only one
participant is designated as the proposer, and hence allowed to issue prepares
and subsequently decrees, at any given time.  In the absence of rejoins, we
are able to provide a very useful system invariant which full Paxos cannot
have:
\begin{quote}
  If a proposer successfully completes a prepare phase, that proposer will
  remain the protocol's designated proposer until it leaves the protocol, by
  failure or otherwise.
\end{quote}
We discuss this invariant further in the following subsection.

\subsection{Support Protocols}

In order to make Paxos efficient and usable, the core protocol needed to be
supplemented with additional protocols which, despite being out-of-band, were
no less integral to the functionality of \Motmot.  We will describe each of
these side protocols in detail.

\begin{description}
  \item[Proposer election:]

  \item[Greet protocol:]

  \item[Out-of-band requests:]

  \item[Reconnection:]

  \item[Commit retry:]

  \item[Log synchronization and truncation:]

  \item[Connection continuations:]
\end{description}

@@

Our simplification to Paxos is a lesson in the importance of end-to-end design
\cite{end2end}.  The \libmotmot library is a backend protocol layer, on top of
which chat semantics are implemented.  Without taking the semantics of the
system's endpoints---the chat clients---into consideration, our consensus
algorithm would have been unnecessarily featureful and complicated.

The importance of out-of-band effort in distributed protocols was also made
very clear to us in our implementation.  Other implementors of Paxos have also
discovered the substantial amount of support needed to make the the protocol
work in practical production-level settings \cite{paxlive}.  Clean designs
and interfaces for these additional protocols was crucial in keeping our
system simple, functional, and understandable.

@@

\thebibliography{7}{
  \bibitem{paxlive}
    T. Chandra, R. Griesemer, and J. Redstone.
    Paxos made live---an engineering perspective.
    In \emph{Proc. 26th Symp. on Principles of Distributed Computing}, pp. 398-407, Aug. 2007.

@@

  \bibitem{rfc}
    P. Resnick, ed.
    Internet Message Format, RFC 2822.
    April 2001.

  \bibitem{end2end}
    J. H. Saltzer, D. P. Reed, and D. D. Clark.
    End-to-end arguments in system design.
    \emph{ACM Transactions on Computer Systems}, 2(4):277-288, Nov. 1984.
}
